---
title: "NCRA Decision Impacts"
author: "Andrew Constable"
date: "2024-09-23"
output: html_document
always_allow_html: true
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library("igraph")

# adjustable arrow heads
source("igraphplot2.R")
environment(plot.igraph2) <- asNamespace('igraph')
environment(igraph.Arrows2) <- asNamespace('igraph')

# functions

fnNetFromMat<-function(eT,Zero.rm=TRUE){
  nodes<-eT[,1]
  row.names(eT)<-nodes
  eT<-eT[,-1] # lose column of names

  edges<-do.call(rbind,
            lapply(row.names(eT),function(r,m){
                r1<-as.vector(unlist(m[r,]))
                if(Zero.rm) p<-which(!is.na(r1) & r1!=0) else p<-which(!is.na(r1))
                c1<-colnames(m)[p]
                res<-data.frame(from=rep(r,length(c1)),to=c1,pressure=r1[p])
                return(res)
                },eT))
  return(list(nodes=nodes,edges=edges))
   }

plotNet<-function(eT,targetNode=NULL,Zero.rm=TRUE,vSize=40,aSize=0.5,eSize=0.5,NfillCols=NULL,NtextCols=NULL,Ecols=NULL){

# Examine the data:

# create data.frame from edge matrix - from, to, pressure
netWork<-fnNetFromMat(eT)  
edges<-netWork$edges
nodes<-netWork$nodes

nodeFillColours<-unlist(sapply(nodes,function(n,Nc){return(if(is.null(Nc)) NA else Nc[[n]] )},NfillCols))
nodeTextColours<-unlist(sapply(nodes,function(n,Nc){return(if(is.null(Nc)) "black" else Nc[[n]] )},NtextCols))

if(!is.null(targetNode)){
 edges<-edges[edges[,1]==targetNode | edges[,2]==targetNode,]  
}
edgeColours<-unlist(sapply(edges[,1],function(n,Ec){return(if(is.null(Ec)) "black" else Ec[[n]])},Ecols))

net <- graph_from_data_frame(d=edges, vertices=nodes, directed=T) 

edgeSize<-edges[,"pressure"]
plot(net,layout=layout_in_circle,edge.curved=0.2,edge.width=edgeSize*eSize,vertex.size=vSize,edge.arrow.size=aSize,vertex.color=nodeFillColours,vertex.frame.color= nodeTextColours,vertex.frame.width=1.5,vertex.label.color=nodeTextColours,edge.color=edgeColours )
} # end plot eT

```
\newpage
## Task 1  
  
### Flow of decisions that result in impacts between systems  
  
Impact Statements Triggered by a Decision  
  
```{r Task-1,echo=FALSE,eval=TRUE}

eTask1_1 <- read.csv("NCRA Data Working File for Modelling(Task 1) 1.1.csv", header=T, as.is=T)
NodeCols<-list(ETF = "#FFCC00",DNS = "#ccff00", CandS = "#80e5ff", PIF="#ffaaee", NE = "#e9afaf", IBE = "#cccccc", HSS = "#afdde9")

EdgeCols<-list(ETF = "#ff6600",DNS = "#668000", CandS = "blue", PIF="#ff00cc", NE = "#501616", IBE = "#1a1a1a", HSS = "#164450")


ArrowSize<-0.5
plotNet(eTask1_1,aSize=ArrowSize,NtextCols=EdgeCols, Ecols=EdgeCols)

Nodes<-eTask1_1[,1]

res<-sapply(Nodes,function(n,eT){
plotNet(eT,targetNode=n,aSize=ArrowSize,NtextCols=EdgeCols, Ecols=EdgeCols)
},eTask1_1)

```

### Flow of decisions resulting in opportunities between systems  
  
Opportunity Statements Triggered by a Decision  
  
```{r Task-2,echo=FALSE,eval=TRUE}

eTask1_2 <- read.csv("NCRA Data Working File for Modelling(Task 1) 1.2.csv", header=T, as.is=T)

ArrowSize<-0.5
plotNet(eTask1_2,aSize=ArrowSize,NtextCols=EdgeCols, Ecols=EdgeCols)

Nodes<-eTask1_2[,1]

res<-sapply(Nodes,function(n,eT){
plotNet(eT,targetNode=n,aSize=ArrowSize,NtextCols=EdgeCols, Ecols=EdgeCols)
},eTask1_2)

```
\newpage
  
### Flow of decisions resulting in impacts between scales  
Impact Statements Triggered by a Decision At What Level  
  
```{r Task-3,echo=FALSE,eval=TRUE}

eTask1_3 <- read.csv("NCRA Data Working File for Modelling(Task 1) 1.3.csv", header=T, as.is=T)

plotNet(eTask1_3,aSize=0.5,Zero.rm=FALSE)

```
  
### Flow of decisions resulting in opportunities between scales  
Opportunity Statements Triggered by a Decision At What Level  
  
```{r Task-4,echo=FALSE,eval=TRUE}

eTask1_4 <- read.csv("NCRA Data Working File for Modelling(Task 1) 1.4.csv", header=T, as.is=T)

plotNet(eTask1_4,aSize=0.3,eSize=0.2,Zero.rm=FALSE)
```


## Sankey plot   

```{r Sankey,eval=TRUE,echo=FALSE}
nodes<-read.csv("Sankey nodes.csv", header=T, as.is=T)
net_1_2<-fnNetFromMat(read.csv("Sankey 1-2.csv", header=T, as.is=T))$edges  
net_2_3<-fnNetFromMat(read.csv("Sankey 2-3.csv", header=T, as.is=T))$edges  
net_3_4<-fnNetFromMat(read.csv("Sankey 3-4.csv", header=T, as.is=T))$edges  
links<-rbind(net_1_2,net_2_3,net_3_4)

# add cols to links with numeric value of nodes
# note source and target need to be zero indexed (hence the inclusion of '-1')

links$source<-as.vector(unlist(sapply(links[,1],function(s,n){which(n==s)},nodes[,1])))-1
links$target<-as.vector(unlist(sapply(links[,2],function(t,n){which(n==t)},nodes[,1])))-1

library(networkD3)
sankeyNetwork(Links = links, Nodes = nodes, Source = "source",Target = "target", Value = "pressure", NodeID = "Name"
                       , fontSize = 10, nodeWidth = 20, iterations=0)

```
