---
title: "NCRA Decision Impacts"
author: "Andrew Constable"
date: "2024-09-23"
output: html_document
always_allow_html: true
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library("igraph")

# adjustable arrow heads
source("igraphplot2.R")
environment(plot.igraph2) <- asNamespace('igraph')
environment(igraph.Arrows2) <- asNamespace('igraph')

# functions

fnNetFromMat<-function(eT,Zero.rm=TRUE){
  nodes<-eT[,1]
  row.names(eT)<-nodes
  eT<-eT[,-1] # lose column of names

  edges<-do.call(rbind,
            lapply(row.names(eT),function(r,m){
                r1<-as.vector(unlist(m[r,]))
                if(Zero.rm) p<-which(!is.na(r1) & r1!=0) else p<-which(!is.na(r1))
                c1<-colnames(m)[p]
                res<-data.frame(from=rep(r,length(c1)),to=c1,pressure=r1[p])
                return(res)
                },eT))
  return(list(nodes=nodes,edges=edges))
   }

fnPointTransTangent<-function(x,y,q,acs,d){  # x = original X, y = original Y
                                         # q = quarter applicable to transformation
                                         # acs = cosine & degrees 
                                         # d = distance along tangent
                 xprime<-ifelse(q==1,d*acs[1,"sin"],d*acs[1,"cos"])
                 yprime<-ifelse(q==1,d*acs[1,"cos"],d*acs[1,"sin"])
                 return(data.frame(x=x+xprime,y=y+yprime))
                 }

fnPointTransAxis<-function(x,y,acs,d){  # x = original X, y = original Y
                                         # q = quarter applicable to transformation
                                         # acs = cosine & degrees 
                                         # d = distance along axis of angle
                 xprime<-d*acs[1,"cos"] #,d*acs[1,"sin"])
                 yprime<-d*acs[1,"sin"] #,d*acs[1,"cos"])
                 return(data.frame(x=x+xprime,y=y+yprime))
                 }

plotNet<-function(eT,targetNode=NULL,Zero.rm=TRUE,vSize=40,aSize=0.5,eSize=0.5,NfillCols=NULL,NtextCols=NULL,Ecols=NULL
                  ,labelVertices="Short",VertexLabelLong=NULL,vLabelLongPosRel=0.2,vLabelLineOffset=1,ScalePlotForLabels=1
                  ,LabelAxisExtend=0.2,TextLineSpace=0.05){

# create data.frame from edge matrix - from, to, pressure
netWork<-fnNetFromMat(eT)  
edges<-netWork$edges
nodes<-netWork$nodes

nodeFillColours<-unlist(sapply(nodes,function(n,Nc){return(if(is.null(Nc)) NA else Nc[[n]] )},NfillCols))
nodeTextColours<-unlist(sapply(nodes,function(n,Nc){return(if(is.null(Nc)) "black" else Nc[[n]] )},NtextCols))
nodeLongLabel<-lapply(nodes,function(n,Nv){return(if(is.null(Nv[[n]])) "" else Nv[[n]] )},VertexLabelLong)

if(!is.null(targetNode)){
 edges<-edges[edges[,1]==targetNode | edges[,2]==targetNode,]  
}
edgeColours<-unlist(sapply(edges[,1],function(n,Ec){return(if(is.null(Ec)) "black" else Ec[[n]])},Ecols))

net <- graph_from_data_frame(d=edges, vertices=nodes, directed=T) 
netLayout<-layout.circle(net)

netLayout<-netLayout*ScalePlotForLabels

edgeSize<-edges[,"pressure"]
if(labelVertices=="Short"){
       plot(net,layout=netLayout,edge.curved=0.2,edge.width=edgeSize*eSize,vertex.size=vSize*ScalePlotForLabels,edge.arrow.size=aSize
            ,vertex.color=NA,vertex.frame.color=nodeTextColours,vertex.frame.width=1.5,vertex.label.cex=ScalePlotForLabels
            ,vertex.label.color=nodeTextColours,edge.color=edgeColours ,rescale=FALSE)
  } else {
       plot(net,layout=netLayout,edge.curved=0.2,edge.width=edgeSize*eSize,vertex.size=vSize*ScalePlotForLabels,edge.arrow.size=aSize
            ,vertex.color=nodeFillColours,vertex.frame.color=nodeTextColours,vertex.frame.width=1.5
            ,vertex.label=NA,edge.color=edgeColours,rescale=FALSE )
  } # end else
if(labelVertices=="Long"){ # add labels outside vertices
      ## Apply labels manually
      #Specify x and y coordinates of labels, adjust outward as desired
      relLabelPos<-vLabelLongPosRel
      x = netLayout[,1]
      y = netLayout[,2]

#create vector of angles for text based on number of nodes (flipping the orientation of the words half way around so none appear upside down)
angleText<- ifelse(atan(-(netLayout[,1]/netLayout[,2]))*(180/pi) < 0,  90 + atan(-(netLayout[,1]/netLayout[,2]))*(180/pi), 270 + atan(-netLayout[,1]/netLayout[,2])*(180/pi))

# restore angles clockwise from horizontal
angleNodeAxis<-angleText
q2<-netLayout[,1]<0 & netLayout[,2]>0
angleNodeAxis[q2]<-angleNodeAxis[q2]-180
q3<-netLayout[,1]<0 & netLayout[,2]<0
angleNodeAxis[q3]<-angleNodeAxis[q3]+180

angle_cos_sin<-data.frame(angle=angleNodeAxis,cos=cos(angleNodeAxis/180*pi),sin=sin(angleNodeAxis/180*pi))
# print(cbind(nodes,netLayout,angleText,angle_cos_sin))

#Apply the text labels with a loop with angle as srt

for (i in 1:nrow(netLayout)) {
  Pt_xy<-fnPointTransAxis(x=netLayout[i,1],y=netLayout[i,2],acs=angle_cos_sin[i,],d=LabelAxisExtend)
  adjX<-ifelse(netLayout[i,1]>=0,0,1); adjY<-0.5
  
  StepDirections<-matrix(c( 1,-1
                          ,-1,-1
                          , 1,-1
                          ,-1,-1),ncol=2,byrow=TRUE,dimnames=list(NULL,c("x","y")))
  
  nLines<-length(nodeLongLabel[[i]])
  tAngle<-(angleNodeAxis[i]%%90)/180*pi
  quad<-(floor(angleNodeAxis[i]/90)+1)
  sd<-StepDirections[quad,]
  
  if(quad==1 | quad==3) trig<-c(sin(tAngle),cos(tAngle)) else trig<-c(cos(tAngle),sin(tAngle))
  oX<-Pt_xy["x"]-(sd["x"]*TextLineSpace*(nLines-1)/2)*trig[1]
  oY<-Pt_xy["y"]-(sd["y"]*TextLineSpace*(nLines-1)/2)*trig[2]
  for(j in 1:nLines){
    
  tX<-oX+sd["x"]*TextLineSpace*(j-1)*trig[1]
  tY<-oY+sd["y"]*TextLineSpace*(j-1)*trig[2]
  
  text(x=tX, y=tY, labels=nodeLongLabel[[i]][j], adj=c(adjX,0.5), pos=NULL, cex=0.9, col=nodeTextColours[i], srt=angleText[i], xpd=T)
  } # end j  

     } # end print labels
} # end long labels
} # end plot eT

```
\newpage
## Task 1  
  
### Flow of decisions that result in impacts between systems  
  
Impact Statements Triggered by a Decision  
  
```{r Task-1,echo=FALSE,eval=TRUE}

eTask1_1 <- read.csv("NCRA Data Working File for Modelling(Task 1) 1.1.csv", header=T, as.is=T)
NodeCols<-list(ETF = "#FFCC00",DNS = "#ccff00", CandS = "#80e5ff", PIF="#ffaaee", NE = "#e9afaf", IBE = "#cccccc", HSS = "#afdde9")
NodeLabelsLong=list(ETF   = c("Economy, Trade","and Finance")
                   ,DNS   = c("Defence and","National Security")
                   ,CandS = c("Communities and","Settlements")
                   ,PIF   = c("Primary Industries","and Food")
                   ,NE    = c("Natural","Environment")
                   ,IBE   = c("Infrastructure and ","Built Environment")
                   ,HSS   = c("Health and","Social Support"))
EdgeCols<-list(ETF = "#ff6600",DNS = "#668000", CandS = "blue", PIF="#ff00cc", NE = "#501616", IBE = "#1a1a1a", HSS = "#164450")



ArrowSize<-0.5
labelVertices<-"Long"
vLabelLongPosRel<-2
ScalePlotForLabels<-0.6
vLabelLineOffset<-0.1

plotNet(eTask1_1,aSize=ArrowSize,NfillCols=EdgeCols,NtextCols=EdgeCols
        ,Ecols=EdgeCols,labelVertices=labelVertices,VertexLabelLong=NodeLabelsLong,vLabelLongPosRel=vLabelLongPosRel,vLabelLineOffset=vLabelLineOffset,ScalePlotForLabels=ScalePlotForLabels,TextLineSpace=0.15)

Nodes<-eTask1_1[,1]


res<-sapply(Nodes,function(n,eT){
#plotNet(eT,targetNode=n,aSize=ArrowSize,NfillCols=EdgeCols,NtextCols=EdgeCols, Ecols=EdgeCols)
plotNet(eT,targetNode=n,aSize=ArrowSize,NfillCols=EdgeCols,NtextCols=EdgeCols
        ,Ecols=EdgeCols,labelVertices=labelVertices,VertexLabelLong=NodeLabelsLong,vLabelLongPosRel=vLabelLongPosRel,vLabelLineOffset=vLabelLineOffset,ScalePlotForLabels=ScalePlotForLabels,TextLineSpace=0.15)
},eTask1_1)

```

### Flow of decisions resulting in opportunities between systems  
  
Opportunity Statements Triggered by a Decision  
  
```{r Task-2,echo=FALSE,eval=TRUE}

eTask1_2 <- read.csv("NCRA Data Working File for Modelling(Task 1) 1.2.csv", header=T, as.is=T)

ArrowSize<-0.5
#plotNet(eTask1_2,aSize=ArrowSize,NtextCols=EdgeCols, Ecols=EdgeCols)
plotNet(eTask1_2,aSize=ArrowSize,NfillCols=EdgeCols,NtextCols=EdgeCols
        ,Ecols=EdgeCols,labelVertices=labelVertices,VertexLabelLong=NodeLabelsLong,vLabelLongPosRel=vLabelLongPosRel,vLabelLineOffset=vLabelLineOffset,ScalePlotForLabels=ScalePlotForLabels,TextLineSpace=0.15)

Nodes<-eTask1_2[,1]

res<-sapply(Nodes,function(n,eT){
#plotNet(eT,targetNode=n,aSize=ArrowSize,NtextCols=EdgeCols, Ecols=EdgeCols)
plotNet(eT,targetNode=n,aSize=ArrowSize,NfillCols=EdgeCols,NtextCols=EdgeCols
        ,Ecols=EdgeCols,labelVertices=labelVertices,VertexLabelLong=NodeLabelsLong,vLabelLongPosRel=vLabelLongPosRel,vLabelLineOffset=vLabelLineOffset,ScalePlotForLabels=ScalePlotForLabels,TextLineSpace=0.15)
},eTask1_2)

```
\newpage
  
### Flow of decisions resulting in impacts between scales  
Impact Statements Triggered by a Decision At What Level  
  
```{r Task-3,echo=FALSE,eval=TRUE}

eTask1_3 <- read.csv("NCRA Data Working File for Modelling(Task 1) 1.3.csv", header=T, as.is=T)

plotNet(eTask1_3,aSize=0.5,Zero.rm=FALSE)

```
  
### Flow of decisions resulting in opportunities between scales  
Opportunity Statements Triggered by a Decision At What Level  
  
```{r Task-4,echo=FALSE,eval=TRUE}

eTask1_4 <- read.csv("NCRA Data Working File for Modelling(Task 1) 1.4.csv", header=T, as.is=T)

plotNet(eTask1_4,aSize=0.3,eSize=0.2,Zero.rm=FALSE)
```


## Sankey plot   

```{r Sankey,eval=TRUE,echo=FALSE}
nodes<-read.csv("Sankey nodes.csv", header=T, as.is=T)
net_1_2<-fnNetFromMat(read.csv("Sankey 1-2.csv", header=T, as.is=T))$edges  
net_2_3<-fnNetFromMat(read.csv("Sankey 2-3.csv", header=T, as.is=T))$edges  
net_3_4<-fnNetFromMat(read.csv("Sankey 3-4.csv", header=T, as.is=T))$edges  
links<-rbind(net_1_2,net_2_3,net_3_4)

# add cols to links with numeric value of nodes
# note source and target need to be zero indexed (hence the inclusion of '-1')

links$source<-as.vector(unlist(sapply(links[,1],function(s,n){which(n==s)},nodes[,1])))-1
links$target<-as.vector(unlist(sapply(links[,2],function(t,n){which(n==t)},nodes[,1])))-1

library(networkD3)
sankeyNetwork(Links = links, Nodes = nodes, Source = "source",Target = "target", Value = "pressure", NodeID = "Name"
                       , fontSize = 10, nodeWidth = 20, iterations=0)

```
