---
title: "NCRA Decision Impacts"
author: "Andrew Constable"
date: "2024-09-15"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library("igraph")

# adjustable arrow heads
source("igraphplot2.R")
environment(plot.igraph2) <- asNamespace('igraph')
environment(igraph.Arrows2) <- asNamespace('igraph')

# functions

plotNet<-function(eT,vSize=40,aSize=1,eSize=0.5,Zero.rm=TRUE){

# Examine the data:
nodes<-eT[,1]
row.names(eT)<-nodes
eT<-eT[,-1] # lose column of names

# create data.frame from edge matrix - from, to, pressure
edges<-do.call(rbind,
            lapply(row.names(eT),function(r,m){
                r1<-as.vector(unlist(m[r,]))
                if(Zero.rm) p<-which(!is.na(r1) & r1!=0) else p<-which(!is.na(r1))
                c1<-colnames(m)[p]
                res<-data.frame(from=rep(r,length(c1)),to=c1,pressure=r1[p])
                return(res)
                },eT))

net <- graph_from_data_frame(d=edges, vertices=nodes, directed=T) 

edgeSize<-edges[,"pressure"]
plot(net,layout=layout_in_circle,edge.curved=0.2,edge.width=edgeSize*eSize,vertex.size=vSize,edge.arrow.size=edgeSize*aSize)
} # end plot eT


```
\newpage
## Task 1  
  
### Task 1.1 Flow of decisions that result in impacts between systems  
  
Impact Statements Triggered by a Decision  
  
```{r Task-1,echo=FALSE,eval=TRUE}

eTask1_1 <- read.csv("NCRA Data Working File for Modelling(Task 1) 1.1.csv", header=T, as.is=T)

plotNet(eTask1_1,aSize=1)
```

### Task 1.2 Flow of decisions resulting in opportunities between systems  
  
Opportunity Statements Triggered by a Decision  
  
```{r Task-2,echo=FALSE,eval=TRUE}

eTask1_2 <- read.csv("NCRA Data Working File for Modelling(Task 1) 1.2.csv", header=T, as.is=T)

plotNet(eTask1_2,aSize=0.2)
```
\newpage
  
### Task 1.3 Flow of decisions resulting in impacts between scales  
Impact Statements Triggered by a Decision At What Level  
  
```{r Task-3,echo=FALSE,eval=TRUE}

eTask1_3 <- read.csv("NCRA Data Working File for Modelling(Task 1) 1.3.csv", header=T, as.is=T)

plotNet(eTask1_3,aSize=0.5,Zero.rm=FALSE)

```
  
### Task 1.4 Flow of decisions resulting in opportunities between scales  
Opportunity Statements Triggered by a Decision At What Level  
  
```{r Task-4,echo=FALSE,eval=TRUE}

eTask1_4 <- read.csv("NCRA Data Working File for Modelling(Task 1) 1.4.csv", header=T, as.is=T)

plotNet(eTask1_4,aSize=0.3,eSize=0.2,Zero.rm=FALSE)
```

## Task 2
### Task 2.1 Flow of decisions resulting in opportunities between scales  
Opportunity Statements Triggered by a Decision At What Level  
  
```{r Task-2_eg,echo=FALSE,eval=TRUE}
library(plotly)
# Eg Sankey code

# Basic
fig <- plot_ly(
    type = "sankey",
    orientation = "h",

    node = list(
      label = c("A1", "A2", "B1", "B2", "C1", "C2"),
      color = c("blue", "blue", "blue", "blue", "blue", "blue"),
      pad = 15,
      thickness = 20,
      line = list(
        color = "black",
        width = 0.5
      )
    ),

    link = list(
      source = c(0,1,0,2,3,3),
      target = c(2,3,3,4,4,5),
      value =  c(8,4,2,8,4,2)
    )
  )
fig <- fig %>% layout(
    title = "Basic Sankey Diagram",
    font = list(
      size = 10
    )
)

fig
###################################
# create canvas

fig <- plot_ly(
    type = "sankey",
    domain = list(
      x =  c(0,1),
      y =  c(0,1)
    ),
    orientation = "h",
    valueformat = ".0f",
    valuesuffix = "TWh"
)
fig <- fig %>% layout(
    title = "Energy forecast for 2050, UK - Department of Energy & Climate Change",
    font = list(
      size = 10
    ),
    xaxis = list(showgrid = F, zeroline = F),
    yaxis = list(showgrid = F, zeroline = F)
  )

# add nodes

library(plotly)
library(rjson)

json_file <- "https://raw.githubusercontent.com/plotly/plotly.js/master/test/image/mocks/sankey_energy.json"
json_data <- fromJSON(paste(readLines(json_file), collapse=""))

fig <- plot_ly(
    type = "sankey",
    domain = list(
      x =  c(0,1),
      y =  c(0,1)
    ),
    orientation = "h",
    valueformat = ".0f",
    valuesuffix = "TWh",

    node = list(
      label = json_data$data[[1]]$node$label,
      color = json_data$data[[1]]$node$color,
      pad = 15,
      thickness = 15,
      line = list(
        color = "black",
        width = 0.5
      )
    )
  ) 
fig <- fig %>% layout(
    title = "Energy forecast for 2050, UK - Department of Energy & Climate Change",
    font = list(
      size = 10
    ),
    xaxis = list(showgrid = F, zeroline = F),
    yaxis = list(showgrid = F, zeroline = F)
)

# add links

library(plotly)
library(rjson)

json_file <- "https://raw.githubusercontent.com/plotly/plotly.js/master/test/image/mocks/sankey_energy.json"
json_data <- fromJSON(paste(readLines(json_file), collapse=""))

fig <- plot_ly(
    type = "sankey",
    domain = list(
      x =  c(0,1),
      y =  c(0,1)
    ),
    orientation = "h",
    valueformat = ".0f",
    valuesuffix = "TWh",

    node = list(
      label = json_data$data[[1]]$node$label,
      color = json_data$data[[1]]$node$color,
      pad = 15,
      thickness = 15,
      line = list(
        color = "black",
        width = 0.5
      )
    ),

    link = list(
      source = json_data$data[[1]]$link$source,
      target = json_data$data[[1]]$link$target,
      value =  json_data$data[[1]]$link$value,
      label =  json_data$data[[1]]$link$label
    )
  ) 
fig <- fig %>% layout(
    title = "Energy forecast for 2050<br>Source: Department of Energy & Climate Change, Tom Counsell via <a href='https://bost.ocks.org/mike/sankey/'>Mike Bostock</a>",
    font = list(
      size = 10
    ),
    xaxis = list(showgrid = F, zeroline = F),
    yaxis = list(showgrid = F, zeroline = F)
)

fig

# define node position

library(plotly)
fig <- plot_ly(
  type = "sankey",
  arrangement = "snap",
  node = list(
    label = c("A", "B", "C", "D", "E", "F"),
    x = c(0.2, 0.1, 0.5, 0.7, 0.3, 0.5),
    y = c(0.7, 0.5, 0.2, 0.4, 0.2, 0.3),
    pad = 10), # 10 Pixel
  link = list(
    source = c(0, 0, 1, 2, 5, 4, 3, 5),
    target = c(5, 3, 4, 3, 0, 2, 2, 3),
    value = c(1, 2, 1, 1, 1, 1, 1, 2)))
fig <- fig %>% layout(title = "Sankey with manually positioned node")

fig




```
