---
title: "Network visualisation & analysis"
author: "Andrew Constable"
date: "2024-09-06"
output: pdf_document
bibliography: Network_refs.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction  
  
This document provides the methodology for developing a network model from base coding in R as well as from an Excel file of relevant interaction matrices.  
  
## Visualisation  
  
Visualisation methods are derived from Ognyanova [-@Ognyanova2023_RN20849].   
  
### Example  
  
```{r EGinput, echo=FALSE, eval=TRUE}
library("igraph")


nEG <- read.csv("Model inputs - test - nodes.csv", header=T, as.is=T)
mEG <- read.csv("Model inputs - test - edges -g.csv", header=T, as.is=T)

# Examine the data:
head(nEG)
head(mEG)
row.names(mEG)<-as.vector(mEG[,1])
mEG<-mEG[,-1] # lose column of names

# winnow nodes and edges to only those nodes that have interactions with other nodes

selfEdge<-(-1) 

res<-as.vector(which((apply(abs(mEG),1,sum) + apply(abs(mEG),2,sum))==2*abs(selfEdge)))
rEGn<-nEG[-res,]
rEGe<-mEG[-res,-res]

# create data.frame from edge matrix - from, to, pressure

edges<-do.call(rbind,
            lapply(row.names(rEGe),function(r,m){
                r1<-as.vector(unlist(m[r,]))
                p<-which(r1!=0)
                c1<-colnames(m)[p]
                res<-data.frame(from=rep(r,length(c1)),to=c1,pressure=r1[p])
                return(res)
                },rEGe))
nodes<-rEGn

# Converting the data to an igraph object:
# The graph_from_data_frame() function takes two data frames: 'd' and 'vertices'.
# 'd' describes the edges of the network - it should start with two columns 
# containing the source and target node IDs for each network tie.
# 'vertices' should start with a column of node IDs.
# Any additional columns in either data frame are interpreted as attributes.

net <- graph_from_data_frame(d=edges, vertices=nodes, directed=T) 

# Examine the resulting object:
class(net)
net 

tkplot(net)

```

## Perturbation analysis  
  
### Qpress installation and application  
  
Qpress is a qualitative network analytical package [@MelbourneThomas2012_RN770] maintained by Simon Wotherspoon on Github (https://github.com/SWotherspoon/QPress?tab=readme-ov-file#readme).  Installation and access to vignettes are detailed in the readme file.  Latest version is 2022.  
  
```{r QpressInstall, echo=FALSE, include=FALSE, eval=FALSE}
# from the readme file
remotes::install_github("SWotherspoon/QPress", build_vignettes = TRUE)
library(QPress)
# help:   ?`QPress-package`
# goto index

vignette("Snowshoe")
vignette("Mesocosm")

tqp<-model.dia("Test 1.dia")

# dia line options for negative influence
#    circle = 8
#    diamond = 5
# dia line options for positive influence
#    line arrow head = 1
#    solid arrow head = 22

# dia node shape
#  ellipse = "Flowchart - Ellipse"
#  rectangle = "Flowchart - Box"
#  rectangle with rounded corners = "Flowchart - Terminal"



```

## References  
  
